Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-05 vesicle_code/docs/adhesion/adhesion.tex
worked on extensional example

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous drop
\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity point) in an extensional flow, $\uu =\chi(-x,y)$,
is a saddle point where two laminar streams converge along an axis and
then diverge in a different direction.  With an active control algorithm
to fix the position of the stagnation point by adjusting the streaming
flow strength with a feedback loop~\cite{Johnson-Chavarria2011_EMJ}, a
particle can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as PIV of
flow inside and around the particle.  To explore the application of such
a fluid trap to measure the adhesion strength between two bound vesicle
membranes, we propose the following thought experiment.

We begin from two vesicles forming a doublet under an adhesion potential
in a quiescent flow as described in Section~\ref{sec:qflow}.  Once a
doublet forms and reaches an equilibrium configuration, we turn on the
fluid trap with the stagnation point placed at the center of the vesicle
doublet.  For simplification we assume that each vesicle in the doublet
has the same reduced area, membrane area, and bending stiffness.  At low
extensional rate, $\chi$, we expect the vesicle doublet to stay fixed at
the fluid trap stagnation point.  On the other hand, the vesicle doublet
may become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate above which
the vesicle doublet cannot stay bound under a given adhesion potential.
The dependence of the critical extensional rate on the adhesion
potential and the mechanical vesicle properties thus provides a means to
probe the physics of membrane adhesion.

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate is $\chi = 0.02$.  The center of the vesicle, denoted with a
  black square, is a stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
  \caption{\label{fig:extensional2} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate is $\chi = 0.07$.  The center of the vesicle, denoted with a
  black square, is a stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area is $A = 0.7$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate is $\chi = 0.1$.  The center of the vesicle, denoted with a black
  square, is a stagnation point.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
  \includegraphics[width = 0.19\textwidth,trim={5cm 2cm 5cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
  \caption{\label{fig:extensional4} Two vesicles in an extensional flow.
  The reduced area is $A = 0.8$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate is $\chi = 0.1$.  Initially, the doublet tilts to an inclination
  angle, and then the extensional rate overcomes the adhesive force and
  the doublet is broken.  The center of the vesicle, denoted with a
  black square, is a stagnation point.}
\end{figure}

We consider six pairs of vesicles with reduced areas 0.70, 0.75, 0.80,
0.85, 0.90, and 0.95, and we vary the extensional rate, $\chi$, between
$10^{-2}$ and $10^{-1}$.  The length of each of the vesicles is fixed,
and the area is changed to achieve the desired reduced area.  Since the
stagnation point can be controlled in an experimental
setting~\cite{Johnson-Chavarria2011_EMJ}, we translate the center of the
doublet at each time step so that the stagnation point occurs exactly in
the middle of the doublet.  With this adjustment, the vesicles either
remain as a doublet in the fluid trap centered around the stagnation
point, or the doublet is broken and the vesicles diverge apart from one
another.

In Figures~\ref{fig:extensional1}--\ref{fig:extensional3}, snapshots of
a fluid trap containing a vesicle doublet of reduced area 0.7 are
simulated at different extensional rates.  While this doublet remains
trapped around the stagnation point, a doublet with a reduced area of
0.8 is broken for a sufficiently large extensional rate
(Figure~\ref{fig:extensional4}).  We summarize the reduced areas and
extensional rates that that result in a fluid trap in left plot of
Figure~\ref{fig:extensionalPhaseDiagram}. Parameter values with a blue
mark result in a fluid trap while the parameter values with a red marks
result in vesicle separation.  For extensional rates and reduced areas
that result in a fluid trap, the initial configuration of the vesicles
have the diverging stream direction parallel to the vesicles' short
axis.  Since the stable configuration of two vesicles in an extensional
flow has the converging direction parallel to the vesicles' short axis,
we expect the vesicles to align themselves with the long axis parallel
to the diverging direction.  We observe in
Figures~\ref{fig:extensional1}--\ref{fig:extensional3} that the vesicles
do in fact rotate, but fall short of aligning their long axis with the
diverging direction.  We summarize the final inclination angle of the
doublet in the left plot of Figure~\ref{fig:extensionalPhaseDiagram}.
The size of the round dots are scaled to the final inclination angle as
defined in the right plot.  At smaller extensional rates, the vesicles
comes closer to aligning their long axis with the diverging direction.
When the doublet is broken at reduced area $0.70$, the extensional rate
is sufficiently large to align the long axis of the vesicle with the
diverging direction and then the vesicles separate.  However, for larger
reduced areas, the separation dynamics is different.  Instead, the
vesicles tilt to an inclination angle and then separate
(Figure~\ref{fig:extensional4}).

\begin{figure}[htp]
  \begin{minipage}{0.63\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \end{minipage}
  \hfill
  \begin{minipage}{0.35\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/angleDefinition.pdf}
    \includegraphics[height=0.45\textwidth]{figs/rotate1.pdf}
    \hfill
    \includegraphics[height=0.45\textwidth]{figs/rotate2.pdf}
  \end{minipage}
  \caption{\label{fig:extensionalPhaseDiagram} {\em Left}: Phase diagram
  of vesicle hydrodynamics in an extensional flow with a fixed Hamaker
  constant $\mathcal{H}=0.7$ and adhesion length scale $\delta = 0.4$.
  For small extensional rates, the doublet structure is not broken (blue
  marks), while at large extensional rates, the doublet is broken (red
  marks).  The size of the round blue marks indicate the final
  inclination angle with the larger dots rotation less ($\theta=\pi$)
  and the smaller dots rotating more ($\theta = \pi/2$).  A square mark
  is used for simulations that did not result in an equilibrium
  inclination angle in the given time horizon.  {\em Right}: The
  definition of the inclination angle.  An inclination angle of $\pi$
  indicates a doublet whose long axis is parallel to the converging
  direction, while an inclination angle of $\pi/2$ indicates a doublet
  whose short axis is parallel to the converging direction.  The square
  mark is a stagnation point while the round point is located at the
  center of the vesicle.}
\end{figure}

\begin{figure}[htp]
  \begin{minipage}{0.45\textwidth}
  \centering
  \includegraphics[height=0.8\textwidth]{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf}
  \end{minipage}
  \hfill
  \begin{minipage}{0.45\textwidth}
    \includegraphics[height=0.8\textwidth]{figs/adR4em1adS7em1_finalInclinationAngle.pdf}
  \end{minipage}
  \caption{\label{fig:extensionalInclinationAngle} {\em Left}: The
  inclination angle of a doublet formed by vesicles of reduced area
  $0.70$ submerged in an extensional flow with an extensional rate of
  $\chi$.  {\em Right}: The final inclination angle of doublets.
  Results are not reported for all the low extensional rates since the
  doublet had not tilted to an equilibrium inclination angle for the
  given time horizon.}
\end{figure}

For very low extensional rates, the time horizon is insufficient for the
doublet to reach an equilibrium state, and these simulations are marked
with a square.  In Figure~\ref{fig:extensionalInclinationAngle}, we plot
the inclination angle as a function of time for four different
extensional rates (left) and the final inclination angle for all
doublets that reach a equilibrium state (right).  We observe that not
only do larger smaller extensional rates result in smaller inclination
angles, but smaller reduced areas also result in smaller inclination
angles.

We further characterize the rotating dynamics of a doublet by examining
the vorticity and total force on the vesicles.  In
Figure~\ref{fig:extensionalVorticity} ...

\begin{figure}[htp]
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity01.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity02.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity03.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity04.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity05.png} \\
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity06.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity07.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity08.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity09.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity10.png}
  \caption{\label{fig:extensionalVorticity} The vorticity and total
  force of two vesicles in a doublet.  The reduced area is $A = 0.7$,
  the extensional rate is $\chi = 0.7$, the adhesion length scale is
  $\delta = 0.4$, and the Hamaker constant is $\mathcal{H} = 0.7$.  When
  the doublet first starts to rotate, the net total force points in the
  direction of the diverging direction.  Next, there is a shearing
  motion between the vesicles that results in a recoil type motion near
  $t=295$.  Finally, the vorticity and the flow between the vesicles
  approaches zero and the result is a fluid trap containing the
  doublet.}
\end{figure}

@@

\begin{figure}[htp]
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image01.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image02.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image03.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image04.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image05.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image06.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image07.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image08.png}
  \caption{\label{fig:doublet090} The formation of a doublet in a shear
  flow.  The reduced area is $0.90$, the Hamaker constant is
  $\mathcal{H}=0.7$,
  the adhesion length scale is $\delta = 0.4$, and the shear rate is
  $\chi=0.5$.}
\end{figure}

\begin{figure}
  \includegraphics[height=0.4\textwidth]{figs/shear_adR4em1Chi1e0_ra-090.pdf}
  \includegraphics[height=0.4\textwidth]{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \caption{{\em Left}: Distance between a pair of vesicles in a planar
  shear flow with shear rate $\chi=0.5$ and adhesion length scale
  $\delta = 0.4$.   The different lines correspond to a linear spacing
  of adhesion strengths ranging from $\mathcal{H}=0.1$ and $\mathcal{H}=1.0$.  {\em Right}:
  Phase diagram of vesicle hydrodynamics in a shear flow.  At a fixed
  reduced area $\triangle A=0.90$ and $\delta =0.4$,  the critical vdW
  potential strength $\mathcal{H}$ for binding of two vesicles in a shear flow
  depends on the shear rate $\chi$.
\label{fig:sflow_phase_diagram}}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in an extensional flow} 
\label{sec:eflow} 
The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  The stagnation
point (zero-velocity point) in an extensional flow, $\uu =\chi(-x,y)$,
is a saddle point where two laminar streams converge along an axis and
then diverge in a different direction.  With an active control algorithm
to fix the position of the stagnation point by adjusting the streaming
flow strength with a feedback loop~\cite{Johnson-Chavarria2011_EMJ}, a
particle can be trapped at the stagnation point for long time scales to
facilitate image acquisition or other detailed measures such as particle
image velocimetry of flow inside and around the particle.  To explore
the application of such a fluid trap to measure the adhesion strength
between two bound vesicle membranes, we propose the following thought
experiment.

We begin with two vesicles identical vesicles forming a doublet under an
adhesion potential in a quiescent flow as described in
Section~\ref{sec:qflow}.  Once a doublet forms and reaches an
equilibrium configuration, we turn on the fluid trap with the stagnation
point placed at the center of the vesicle doublet.  At low extensional
rates, $\chi$, we expect the vesicle doublet to stay fixed at the fluid
trap stagnation point.  On the other hand, the vesicle doublet may
become unstable and eventually separate at higher extensional rates.
Thus we expect there to exist a critical extensional rate above which
the vesicle doublet cannot stay bound under a given adhesion potential.
Therefore, the dependence of the critical extensional rate on the
adhesion potential and the mechanical vesicle properties provides a
means to probe the physics of membrane adhesion.

\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}
  \caption{\label{fig:extensional1} A doublet inside a fluid trap.  The
  reduced area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate is $\chi = 0.02$.  The center of the vesicle, denoted with a
  black square, is a stagnation point.}

  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png}
  \caption{\label{fig:extensional2} A doublet inside a fluid trap.  The
  reduced area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H}
  = 0.7$, the adhesion length scale is $\delta = 0.4$, and the
  extensional rate is $\chi = 0.07$.  The center of the vesicle, denoted
  with a black square, is a stagnation point.}

  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional3} Two vesicles in an extensional flow.
  The reduced area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} =
  0.7$, the adhesion length scale is $\delta = 0.4$, and the extensional
  rate is $\chi = 0.1$.  The center of the vesicle, denoted with a black
  square, is a stagnation point.}

  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
  \caption{\label{fig:extensional4} A doublet broken by an extensional
  flow.  The reduced area is $\Delta A = 0.8$, the Hamaker constant is
  $\mathcal{H} = 0.7$, the adhesion length scale is $\delta = 0.4$, and
  the extensional rate is $\chi = 0.1$.  Initially, the doublet tilts to
  an inclination angle, and then the extensional rate overcomes the
  adhesive force and the doublet is broken.  The center of the vesicle,
  denoted with a black square, is a stagnation point.}
\end{figure}

We consider vesicle doublets with reduced areas 0.70, 0.75, 0.80, 0.85,
0.90, and 0.95, all with a length of $\pi$, and we vary the extensional
rate, $\chi$, between $10^{-2}$ and $10^{-1}$.  Since the stagnation
point can be controlled in an experimental
setting~\cite{Johnson-Chavarria2011_EMJ}, we translate the center of the
doublet at each time step so that the stagnation point occurs exactly in
the middle of the doublet.  With this adjustment, the vesicles either
remain as a doublet in the fluid trap centered around the stagnation
point, or the doublet is broken and the vesicles separate from one
another.  In Figures~\ref{fig:extensional1}--\ref{fig:extensional3},
snapshots of a fluid trap containing a vesicle doublet of reduced area
$\Delta A = 0.7$ are simulated at different extensional rates.  While
this doublet remains trapped around the stagnation point for all the
considered extensional rates, a doublet with a reduced area of $\Delta A
= 0.8$ is broken with an extensional rate $\chi = 0.1$
(Figure~\ref{fig:extensional4}).  We summarize the reduced areas and
extensional rates that result in a fluid trap in left plot of
Figure~\ref{fig:extensionalPhaseDiagram}. Parameter values with a blue
mark result in a fluid trap and parameter values with a red mark result
in vesicle separation.

For extensional rates and reduced areas that result in a fluid trap, the
initial configuration of the vesicles have the diverging stream
direction parallel to the vesicles' short axis.  Since the stable
configuration of two vesicles in an extensional flow has the converging
direction parallel to the vesicles' short axis, we expect the vesicles
to align themselves with the long axis parallel to the diverging
direction.  We observe in
Figures~\ref{fig:extensional1}--\ref{fig:extensional3} that the vesicles
do in fact rotate, but fall short of aligning their long axis with the
diverging direction.  We summarize the final inclination angle of the
doublet in the left plot of Figure~\ref{fig:extensionalPhaseDiagram}.
The size of the round dots are scaled to the final inclination angle as
defined in the right plot.  At smaller extensional rates, the vesicles
come closer to aligning their long axis with the diverging direction.
When the doublet is broken at reduced area $\Delta A = 0.80$, the
extensional rate is sufficiently large to align the long axis of the
vesicle with the diverging direction and then the vesicles separate
(Figure~\ref{fig:extensional4}).  For very low extensional rates, the
time horizon is insufficient for the doublet to reach an equilibrium
state, and these simulations are marked with a square.

\begin{figure}[htp]
  \begin{minipage}{0.53\textwidth}
    \includegraphics[width=0.33\textwidth]{figs/angleDefinition.pdf} \:
    \includegraphics[width=0.34\textwidth]{figs/rotate1.pdf} \:
    \includegraphics[width=0.26\textwidth]{figs/rotate2.pdf}
  \end{minipage}
  \hfill
  \begin{minipage}{0.45\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \end{minipage}
  \caption{\label{fig:extensionalPhaseDiagram} {\em Left}: The
  definition of the inclination angle.  An inclination angle of $\pi$
  indicates a doublet whose long axis is parallel to the diverging
  direction, while an inclination angle of $\pi/2$ indicates a doublet
  whose short axis is parallel to the diverging direction.  The square
  mark is a stagnation point and the round point is located at the
  center of the vesicle. {\em Right}: A phase diagram of the long-time
  behavior of a fluid trap formed by two adhering vesicles in an
  extensional flow.  The Hamaker constant is $\mathcal{H}=0.7$ and the
  adhesion length scale is $\delta = 0.4$.  For small extensional rates,
  the doublet structure is not broken (blue marks), while at large
  extensional rates, the doublet is broken (red marks).  The size of the
  round blue marks indicates the final inclination angle with the larger
  dots rotating less ($\theta=\pi$) and the smaller dots rotating more
  ($\theta = \pi/2$).  A square mark is used for simulations that did
  not result in an equilibrium inclination angle in the given time
  horizon.}
\end{figure}

In Figure~\ref{fig:extensionalInclinationAngle}, we plot the inclination
angle of the doublet as a function of time for four different
extensional rates (left) and the final inclination angle for all
doublets that reach an equilibrium state (right).  We observe that not
only do larger smaller extensional rates result in smaller inclination
angles, but smaller reduced areas also result in smaller inclination
angles.  Moreover, the critical Hamaker constant that determines if the
doublet remains trapped or is broken decreases with an increasing
extensional rate.

\begin{figure}[htp]
  \includegraphics[height=0.35\textwidth]{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf}
  \includegraphics[height=0.35\textwidth]{figs/adR4em1adS7em1_finalInclinationAngle.pdf}
  \caption{\label{fig:extensionalInclinationAngle} {\em Left}: The
  inclination angle of a doublet formed by vesicles of reduced area
  $\Delta A = 0.7$ submerged in an extensional flow with an extensional
  rate of $\chi$.  {\em Right}: The final inclination angle of doublets.
  At larger extensional rates, the doublet is broken by the flow.
  Results are not reported for some of the low extensional rates since
  the doublet had not tilted to an equilibrium inclination angle for the
  given time horizon.}
\end{figure}

Finally, to further characterize the rotating dynamics of a doublet, we
compute and plot the vorticity and total force on the vesicles in
Figure~\ref{fig:extensionalVorticity}. We consider the doublet formed
with a two vesicles of reduced area $\Delta A = 0.7$, adhesion length
scale $\delta = 0.4$, Hamaker constant $\mathcal{H} = 0.7$, and
extensional rate $\chi = 0.7$.  The doublet's inclination angle remains
close to $\pi$ for a period of time (see the left plot of
Figure~\ref{fig:extensionalInclinationAngle}).  Then, as the doublet
starts to rotate, the total force first points in the direction of the
diverging direction and then rotates until it is parallel with the
vesicles' long axis.  This results in a shearing motion between the
vesicles that results in a recoil type motion near $t=295$.  Finally,
the vorticity and the flow between the vesicles approaches zero and the
result is a fluid trap containing the doublet.
\begin{figure}[htp]
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity01.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity02.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity03.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity04.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity05.png} \\
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity06.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity07.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity08.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity09.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity10.png}
  \caption{\label{fig:extensionalVorticity} The vorticity and total
  force of two vesicles in a doublet.  The reduced area is $\Delta A = 0.7$,
  the extensional rate is $\chi = 0.7$, the adhesion length scale is
  $\delta = 0.4$, and the Hamaker constant is $\mathcal{H} = 0.7$.}
\end{figure}

@@

\begin{figure}[htp]
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image01.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image02.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image03.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image04.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image05.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image06.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image07.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image08.png}
  \caption{\label{fig:doublet090} The formation of a doublet in a shear
  flow.  The reduced area is $\Delta A = 0.90$, the Hamaker constant is
  $\mathcal{H}=0.7$, the adhesion length scale is $\delta = 0.4$, and
  the shear rate is $\chi=0.5$.}
\end{figure}

\begin{figure}
  \includegraphics[height=0.4\textwidth]{figs/shear_adR4em1Chi1e0_ra-090.pdf}
  \includegraphics[height=0.4\textwidth]{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \caption{{\em Left}: Distance between a pair of vesicles in a planar
  shear flow with shear rate $\chi=0.5$ and adhesion length scale
  $\delta = 0.4$.   The different lines correspond to a linear spacing
  of adhesion strengths ranging from $\mathcal{H}=0.1$ and
  $\mathcal{H}=1.0$.  {\em Right}: Phase diagram of vesicle
  hydrodynamics in a shear flow.  At a fixed reduced area $\Delta
  A=0.90$ and $\delta =0.4$,  the critical vdW potential strength
  $\mathcal{H}$ for binding of two vesicles in a shear flow depends on
  the shear rate $\chi$.
\label{fig:sflow_phase_diagram}}
\end{figure}
